Probability of causal contact between interstellar civilizations through MonteCarlo simulations

### 2019-07-31 manuscript.tex
un poco mas de texto

A common approach to discuss the possible number of extraterrestrial
intelligent civilizations (ETI) has been the use of the Drake equation
\citep{Gleiser2010, Prantzos2013, Haqq-Misra2017}.
%
However, the uncertainties in the factors make it less prone to a
formal application to define searching strategies or to compute
actual numbers, but is used to organize the discussion instead.
%
Some modifications to the original idea of the equation have tried to
imprint a stochastic nature, or to propose a probabilistic approach,
or to consider the temporal structure which is missing in the
equation.

%
Temporal aspects of the distribution of communicating civilizations
and their contacts have also been explored by several authors
\citep{Fogg1987, Forgan2011, Balbi2018},
%The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
%Temporal dispersion of the emergence of intelligence: an inter-arrival time analysis
%The spatiotemporal aspects of SETI
%
as well as efforts on considering the stochastic nature of the Drake equation
\citep{Glade2011}.
%
The simulation approach has also benn considered
\citep{Forgan2008, Forgan2010}.
%New numerical determination of habitability in the Galaxy: the SETI connection
%
Since the factors in the Drake equation are uncertain, we propose to
avoid the frequentist approach of this equation
and to explore a parameter space, where instead of computing a final
number, we provide a statistical distribution that gives conditional
probabilities.
%
The only fact that can be stated with certainty is that for the number
of years SETI projects have been working we have not received any
signal within the conditions stablished by SETI.

@@

%Our results highlight the necessity, which has been noted
%elsewhere (Ćirković, 2004), of modeling 
%
%
%While the Drake equation remains a very useful
%guide in identifying and analyzing the various factors in-
%volved in the problem and their relative probabilities, our
%approach could provide a better framework with which to
%perform a statistical analysis when taking into account evo-
%lutionary processes. In particular, we have shown that the
%Drake equation can misestimate the number of detectable
%communicating species if spatial and temporal dependencies
%are important. This should suggest some caution in the un-
%critical application of such an equation. In this paper, we offer
%an exemplificative illustration of the effect of different
%probability distributions for the typical timescales involved
%in the problem. Ideally, it would be desirable to arrive at a
%mathematical description of such distributions based on a
%modeling of the underlying astrophysical and biological
%processes. This points out a possible future direction of in-
%vestigation in which an integrated evolutionary (and coevo-
%lutionary) approach to the problem would be a central focus
%of interest (for more on this perspective, see, e.g., Chaisson

@@

Some complex stochastic processes can be efficiently modeled with the
discrete--event (DE) simulation approach.
%
Simulations are suitable tools to analyze systems that evolve with
time and involve randomness.
%
In general, simulations require less assumptions and simplifications
than theoretical approaches, and can be applied to systems where a
theoretical model can not be found.
%
A system described with the DE paradigm is characterized by a set of
actors and events, where actors interact causally through a series of
events on a time line and process these events in chronological order
\citep{ptolemaeus_system_2014, chung_simulation_2003,
simulation_ross_2012}.
%
On the particular case of the difusion of signals of intelligence in
the Galaxy, the discrete events method can be performed taking into
account a small number of variables.
%
The temporal structure is defined by two distribution parameters that
represent the mean time interval that an intelligent civilization can
emit and receive signals, and the mean time interval between the
emergence of an intelligent comunicating civilization and the next
one.
%
The spatial structure of the simulation is given by the size and shape
of the Galactic Habitable Zone and the maximum distance a signal can
travel to be detected.
%
The parameters for the temporal distributions also determine the
spatial properties.
%
For example, the density of active cetis in the Galaxy depend on these
two parameters.
%
Also, some hypothesis must be made in order to complete the
simulation.
%
Firstly, we assume that the distribution of the appeareance of new
CETIs is a stationary Poisson distribution.
%
Regarding the duration of a CETI, we propose that the distribution of
the duration of CETIs is a stationary exponential distribution.
%
\citet{Balbi2018} stress the fact that it would be desirable to arrive
at a theoretical statistical distribution of the lifespan of
civilizations, preferentially on the basis of the underlying
astrophysical and biological processes.
%
\citet{Maccone} argues that this distribution should be a log normal.
(EXPAND)

@@

\setlength{\tabcolsep}{10pt}
\begin{table}
\centering
\begin{tabular}{cl}
\hline
   $\tau_{a}$ & Mean temporal separation between consecutive awakenings \\
   $\tau_{s}$ & Mean lifetime of a CETI \\
   D$_{max}$ & Maximum reach of a message \\
\hline
\end{tabular}
\caption{Definition of variables that are part of the simulation.}
\label{T_simu}
\end{table}

@@

\subsection{Age distribution of civilizations}

\subsection{Rate of appeareance of civilizations}

>>>>

A common approach to discuss the possible number of extraterrestrial
intelligent civilizations (ETI) has been the use of the Drake equation
\citep{Gleiser2010, Prantzos2013, Haqq-Misra2017}.
%
However, the uncertainties in the factors make it less prone to a
formal application to define searching strategies or to compute
actual numbers, but is used to organize the discussion instead.
%
Some modifications to the original idea of the equation have tried to
imprint a stochastic nature, or to propose a probabilistic approach,
or to consider the temporal structure which is missing in the
equation.
%
Temporal aspects of the distribution of communicating civilizations
and their contacts have also been explored by several authors
\citep{Fogg1987, Forgan2011, Balbi2018},
%The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
%Temporal dispersion of the emergence of intelligence: an inter-arrival time analysis
%The spatiotemporal aspects of SETI
%
as well as efforts on considering the stochastic nature of the Drake equation
\citep{Glade2011}.
%
The simulation approach has also benn considered
\citep{Forgan2008, Forgan2010}.
%New numerical determination of habitability in the Galaxy: the SETI connection
%
Since the factors in the Drake equation are uncertain, we propose to
avoid the frequentist approach of this equation
and to explore a parameter space, where instead of computing a final
number, we provide a statistical distribution that gives conditional
probabilities.
%
The only fact that can be stated with certainty is that for the number
of years SETI projects have been working we have not received any
signal within the conditions stablished by SETI.

@@

%While the Drake equation remains a very useful
%guide in identifying and analyzing the various factors in-
%volved in the problem and their relative probabilities, our
%approach could provide a better framework with which to
%perform a statistical analysis when taking into account evo-
%lutionary processes. In particular, we have shown that the
%Drake equation can misestimate the number of detectable
%communicating species if spatial and temporal dependencies
%are important. This should suggest some caution in the un-
%critical application of such an equation. In this paper, we offer
%an exemplificative illustration of the effect of different
%probability distributions for the typical timescales involved
%in the problem. Ideally, it would be desirable to arrive at a
%mathematical description of such distributions based on a
%modeling of the underlying astrophysical and biological
%processes. This points out a possible future direction of in-
%vestigation in which an integrated evolutionary (and coevo-
%lutionary) approach to the problem would be a central focus
%of interest (for more on this perspective, see, e.g., Chaisson

@@

Some complex stochastic processes can be efficiently modeled with the
discrete--event (DE) simulation approach.
%
Simulations are suitable tools to analyze systems that evolve with
time and involve randomness.
%
In general, simulations require less assumptions and simplifications
than theoretical approaches, and can be applied to systems where a
theoretical model can not be found.
%
A system described with the DE paradigm is characterized by a set of
actors and events, where actors interact causally through a series of
events on a time line and process these events in chronological order
\citep{ptolemaeus_system_2014, chung_simulation_2003,
simulation_ross_2012}.
%
This method is well suited for the 
particular case of the difusion of intelligence signals in
the Galaxy.
%
Assuming some simple hypotheses, 
the discrete events method can be performed taking into
account a small number of variables, which allows to analyze the
results as a function of those variables.

In what follows, we describe the experimental setup chosen to estimate
the probabilities of causal contacts and several derived quantities in
terms of three independent parameters, namely, the mean time span
between the appeareance of consecutive communicating civilizations
($\tau_a$), their mean lifetime ($\tau_s$), and the maximum distance a
signal can be detected by another civilization ($D_{max}$).
%
Intuitively, the shorter $\tau_a$ and the larger $\tau_s$ and
$D_{max}$, there is a greater probability of the existence of causal
contacts between pairs of intelligent civilizations in the Galaxy
(Table \ref{T_simu}).
%
The causal connection can be produced without the need of the two
civilizations being concurrent or active at the same time.
%
Each CETI will fill an annular region of the Galaxy, limited by two
surfaces.
%
The leading front, or surface of first contact (SFC) grows from the
central CETI until it reaches the distance $D_{max}$.
%
The trailing front, or surface of last contact (SLC) also grows from
the central CETI, with a delay with respect to the SFC equivalent to
the lifetime of the CETI, and produces an annular region that shrinks
until it reaches $D_{max}$, when all signal from the central CETI is
lost.
%
In the 2D simulation only the intersection of these spheres with the
plane of the galaxy is relevant, and produce the corresponding
circumferences.

The temporal structure of the process is defined by two distribution parameters that
represent the mean time interval that an intelligent civilization can
emit and receive signals, and the mean time interval between the
emergence of an intelligent comunicating civilization and the next
one.
%
The spatial structure of the simulation is given by the size and shape
of the Galactic Habitable Zone and the maximum distance a signal can
travel to be detected.
%
The parameters for the temporal distributions also determine the
spatial properties since the density of active cetis in the Galaxy depend on these
two parameters.
%
For example, a small $\tau_a$ and a large $\tau_s$ will produce a
Galaxy densely populated woth CETIs, and viceversa.
%
Also, some hypotheses must be made in order to complete the
simulation.
%
Firstly, we assume that the distribution of the appeareance of new
CETIs is a stationary Poisson distribution.
%
Regarding the duration of a CETI, we propose that the distribution of
the duration of CETIs is a stationary exponential distribution.
%
\citet{Balbi2018} stress the fact that it would be desirable to arrive
at a theoretical statistical distribution of the lifespan of
civilizations, preferentially on the basis of the underlying
astrophysical and biological processes.
%
\citet{Maccone} argues that this distribution should be a log normal.
(EXPAND)

@@

\setlength{\tabcolsep}{10pt}
\begin{table}
\centering
\begin{tabular}{cl}
\hline
   \multicolumn{2}{l}{independent variables} \\
\hline
   $\tau_{a}$ & Mean temporal separation between\\
              & consecutive awakenings \\
   $\tau_{s}$ & Mean lifetime of a CETI \\
   D$_{max}$ & Maximum reach of a message \\
\hline
\multicolumn{2}{l}{fixed variables and assumptions} \\
\hline
   & All CETIs share the same statistical properties\\
   & The process is homogeneous \\
   & The scan of the sly is fully efficient\\
   & Absence of panspermia or colonization \\
   & The GHZ is a 2--dimensional ring \\
20000ly    & Inner radius of the GHZ \\
60000ly    & Outer radius of the GHZ \\
1.e6 y    & Time span of the simulation \\

\hline
\end{tabular}
\caption{Definition of independent variables and fixed paramenters 
   that are part of the simulation.}
\label{T_simu}
\end{table}

@@
